---
title: "Cortisol in matched helper and trapped pigs"
author: "Liza Moscovice, Anja Eggert and Jean-Loup Rault"
date: "`r Sys.Date()`" 
editor: visual
code-fold: false
toc: true
format: html
self-contained: true
---

# Libraries

```{r, libraries, warning=FALSE, message=FALSE, echo=TRUE}
library(tidyverse)     # tidy universe
library(viridis)       # color scale
library(rstatix)       # get_summary_stats()
library(ggpubr)        # paired ggplot
```

```{r, my_theme, echo = FALSE}
my_theme = theme_classic() +
  theme(axis.title = element_text(face="bold", size=14),
        axis.text  = element_text(size=12, angle = 0, vjust = 0.5),
        plot.title = element_text(face="bold", size=14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

```{r, seed}
set.seed(1989)
```

# Data

Comparison of post-release salivary cortisol in trapped pigs and their helpers for a subset of n = 50 trials.

## Read data

```{r, data}
dat <- read_csv("../data/cortisol-helper-trapped.csv")
```

## Data summary

```{r, data-sum}
dat %>% 
  rstatix::get_summary_stats(c(cort_potential_helper, cort_trapped), 
                             type = "common")
```

# Paired t-test

```{r, t-test}
t.test(dat$cort_trapped, dat$cort_potential_helper,
       paired = TRUE, alternative = "two.sided")
```

# Pearson correlation

```{r, corr}
cor.test(dat$cort_trapped, dat$cort_potential_helper,
         method = "pearson")
```

# Plot

```{r, plot-1}
plot <- dat %>% 
  ggpaired(
    cond1      = "cort_potential_helper",
    cond2      = "cort_trapped",
    color      = "condition",
    palette    = c("#377EB8","#C51B7D"),
    line.color = "gray84",
    line.size  = 0.5,
    point.size = 3,
    alpha      = 0.8) +
  scale_x_discrete(labels = c("Helper", "Trapped pig")) + 
  scale_y_continuous(limits = c(0, 14), breaks = seq(0, 14, by = 2)) +
  labs(x = "Paired subjects",
       y = "Post-release cortisol (ng/mL)") + 
  my_theme +
  theme(legend.position = "none")
```

```{r, plot-2, fig.height=8, fig.width=10, warning=FALSE}
plot
```

# How to cite R

"All analyses were performed using R Statistical Software (version 4.2.0; R Core Team 2022)".

Reference: R Core Team (2022). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

```{r, cite-r}
citation()
version$version.string
```

```{r, cite-packages}
citation("tidyverse")
citation("viridis")
citation("ggpubr")
```

# Session Info

```{r, session}
sessionInfo()
```
