---
title: "Group latencies to open compartments over familiarization days"
author: "Liza Moscovice, Anja Eggert, Christian Manteuffel and Jean-Loup Rault"
date: "`r Sys.Date()`" 
editor: visual
code-fold: false
toc: true
format: html
self-contained: true
---

# Libraries

```{r, libraries, warning=FALSE, message=FALSE, echo=TRUE}
library(tidyverse)     # tidy universe
library(lmerTest)      # mixed model lme4 with p-values
library(emmeans)       # emmeans
```

```{r, my_theme, echo = FALSE}
my_theme = theme_classic() +
  theme(axis.title = element_text(face = "bold", size = 14),
        axis.text  = element_text(size = 12, angle = 0, vjust = 0.5),
        plot.title = element_text(face = "bold", size = 14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

```{r, seed}
set.seed(1989)
```

# Data

## Read data

```{r, data, warning = FALSE}
dat <- read_csv("../data/Fig-S3-latency-familiarization.csv")
```

## Adjust data types

```{r, data-adj}
dat <- dat %>% 
  mutate_at(vars(familiarization_day,
                 location.box,
                 group.daily.test.order), ~as.factor(.))
```

## Data summary

```{r, data-sum}
dat %>% 
  group_by(familiarization_day, location.box)%>%
  rstatix::get_summary_stats(c(latency.open.min), 
                             show = c("mean", "sd", 
                                      "min", "max", 
                                      "median", "iqr")) %>% 
  ungroup()
```

# Linear mixed model

## Run full model

```{r, mod-cont}
contr = lmerControl(optCtrl = list(maxfun = 10000000),
                    calc.derivs = FALSE)
```

```{r, mod}
mod <- lmer(log(latency.open.min) ~
              # test predictors
              location.box * familiarization_day + 
              # control predictor
              group.daily.test.order + 
              # random/nested factor
              (1|unique.group),
            # data set
            data = dat,
            REML = TRUE,
            # numerical optimizer
            control = contr)
```

## Run reduced model

```{r, mod-red}
mod.red <- lmer(log(latency.open.min) ~
                  # no test predictors
                  # control predictor
                  group.daily.test.order + 
                  # random/nested factor
                  (1|unique.group),
                # data set
                data = dat,
                REML = TRUE,
                # numerical optimizer
                control = contr)
```

## Model comparison

```{r, mod-anova}
anova(update(mod.red, REML = FALSE), 
      update(mod, REML = FALSE), 
      test = "Chisq")
```

## Summary of model

```{r, mod-sum-1}
summary(mod)
```

```{r, mod-sum-2}
round(drop1(mod, test = 'Chisq'), 3)
as.data.frame(round(confint(object = mod), 3)) 
```

## Performance of model

```{r, mod-perf, fig.height=25, fig.width=10, warning=FALSE }
performance::check_model(mod)
```

# Estimated Marginal Means and multiple comparisons

## Within comparisons

```{r, mod-emm}
emm <- emmeans(mod, 
               specs  = pairwise ~ familiarization_day | location.box,
               type   = "response",
               adjust = "tukey")

emm
```

# Plot

## Compact letter display

```{r, mod-cld}
cld <- multcomp::cld(emm$emmeans, Letters=letters)
cld$.group <- str_replace_all(cld$.group, fixed(" "), "")

cld <- as_tibble(cld) %>% 
  arrange(familiarization_day,
          location.box) %>% 
  # only letters next to "side compartment"
  filter(location.box == "Side")

cld
```

## Plot with emmeans

```{r, plot-1}
# prepare x-position of emmeans (only side compartment)
cld$xpos   <- c(1.3, 2.3, 3.3, 4.3, 5.3)
cld$xpos.l <- c(1.4, 2.4, 3.4, 4.4, 5.4)

plot <- dat %>%
  ggplot() +
  geom_boxplot(aes(x   = familiarization_day, 
                   y   = latency.open.min,
                   col = location.box), 
               outlier.shape = NA, width = 0.5, show.legend = FALSE) +
  geom_jitter(aes(x     = familiarization_day, 
                  y     = latency.open.min, 
                  col   = location.box,
                  shape = location.box),
              position = position_jitterdodge(jitter.width = 0.1,
                                              dodge.width  = 0.5),
              size = 3, alpha = 0.7) +
  scale_colour_manual(labels = c("Front compartment",
                                 "Side compartment"),
                      values = c("royalblue1", "royalblue4"))+
  scale_shape_manual(labels = c("Front compartment",
                                "Side compartment"),
                     values = c(17, 19)) +
  scale_y_continuous(lim = c(0, 8), breaks = seq(0, 8, 2)) +
  scale_x_discrete() +
  geom_point(data=cld, col="black", size=3,
             aes(x=xpos, y=response)) +
  geom_errorbar(data=cld, col="black", width=0.08,
                aes(x=xpos,
                    ymin=response-SE, ymax=response+SE)) +
  geom_text(data=cld, col="black",
            aes(y=response, x=xpos.l, label=.group, hjust = 0)) +
  labs(x = "Familiarization day",
       y = "Latency to first open door (min)") +
  my_theme +
  theme(legend.justification = c(1.0, 0.9),
        legend.position      = c(1.0, 0.9),
        legend.background = element_rect(fill     = NA,
                                         colour   = "black",
                                         size     = 0.3,
                                         linetype = "solid"),
        legend.title = element_blank(),
        legend.text = element_text(size = 12)) +
  guides(shape = guide_legend(override.aes = list(alpha = 1,
                                                  size  = 4)))
```

```{r, plot-2, fig.height=10, fig.width=10, warning=FALSE}
plot
```

# How to cite R

"All analyses were performed using R Statistical Software (version 4.2.0; R Core Team 2022)".

Reference: R Core Team (2022). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

```{r, cite-r}
citation()
version$version.string
```

```{r, cite-packages}
citation("tidyverse")
citation("lmerTest")
citation("emmeans")
```

# Session Info

```{r, session}
sessionInfo()
```
