---
title: "Latency across trials"
author: "Liza Moscovice, Anja Eggert and Jean-Loup Rault"
date: "`r Sys.Date()`" 
editor: visual
code-fold: false
toc: true
format: html
---

# Libraries

```{r, libraries, warning=FALSE, message=FALSE, echo=TRUE}
library(tidyverse)     # tidy universe
library(viridis)       # color scale
```

```{r, my_theme, echo = FALSE}
my_theme = theme_classic() +
  theme(axis.title = element_text(face="bold", size=14),
        axis.text  = element_text(size=12, angle = 0, vjust = 0.5),
        plot.title = element_text(face="bold", size=14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

```{r, seed}
set.seed(1989)
```

# Data

Latency for 75 helper pigs to release trapped pigs across 10 trials within each group.

```{r, data}
dat <- read_csv("../data/latency-trials.csv")

str(dat)
```

# Data summary

-   How many pigs were released by another helper pig and how many by a human?

```{r, data-1}
table(dat$helper)
```

-   What was the overall latency time?

```{r, data-2}
dat %>%
  filter(helper == "Pig") %>%
  rstatix::get_summary_stats(lat.helped.min, 
                             type = "common")
```

-   What was the latency time for each test day?

```{r, data-3}
dat %>%
  filter(helper == "Pig") %>%
  group_by(trapped.order) %>% 
  rstatix::get_summary_stats(lat.helped.min, 
                             type = "common")
```

# Plot

```{r, plot-1, fig.height=7, fig.width=10, warning=FALSE}
plot <- dat %>% 
  ggplot() + 
  geom_boxplot(data = subset(dat, helper == "Pig"),
               aes(x = as.factor(trapped.order), 
                   y = lat.helped.min),
               outlier.shape = NA, width = 0.5) + 
  geom_jitter(aes(x     = as.factor(trapped.order), 
                  y     = lat.helped.min,
                  col   = id.group,
                  shape = helper), 
              size = 3, width = 0.15, alpha = 0.8) +
  scale_shape_manual(values = c(17, 19),
                     labels = c("=  pig released by human", "")) +
  scale_color_manual(values = c(viridis(n=8)),
                     guide="none") +
  scale_y_continuous(lim = c(0, 20.01), breaks = seq(0, 20, 5)) +
  labs(x = "Group testing order",
       y = "Latency for helper to release trapped pig (min)",
       shape = "") +
  my_theme +
  guides(shape = guide_legend(override.aes = list(size  = 4,
                                                  shape = c(17, NA)))) +
  theme(legend.justification = c(1,0.6),
        legend.position = c(1,0.6),
        legend.background = element_rect(fill = NA, colour ="black",
                                         size=0.3, linetype="solid"),
        legend.text = element_text(size=12))
```

```{r, plot-2, fig.height=10, fig.width=10, warning=FALSE}
plot
```

# How to cite R

"All analyses were performed using R Statistical Software (version 4.2.0; R Core Team 2022)".

Reference: R Core Team (2022). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

```{r, citations}
citation("tidyverse")
citation("viridis")
```

# Session Info

```{r, session}
sessionInfo()
```
