---
title: "Proportion of pigs that open doorduring familiarization"
author: "Liza Moscovice, Anja Eggert and Jean-Loup Rault"
date: "`r Sys.Date()`" 
editor: visual
code-fold: false
toc: true
format: html
---

# Libraries

```{r, libraries, warning=FALSE, message=FALSE, echo=TRUE}
library(tidyverse)     # tidy universe
```

```{r, my_theme, echo = FALSE}
my_theme = theme_classic() +
  theme(axis.title = element_text(face="bold", size=14),
        axis.text  = element_text(size=12, angle = 0, vjust = 0.5),
        plot.title = element_text(face="bold", size=14),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

```{r, seed}
set.seed(1989)
```

# Data

This data table contains the proportion of pigs that opened the door across 5 training days and the cumulative score.

## Read data

```{r, data, warning = FALSE}
dat <- read_csv("../data/proportion-familiarization.csv")
```

## Adjust data types

```{r}
dat <- dat %>%
  mutate_at(vars(familiarization_day,
                 level), ~as.factor(.)) %>% 
  mutate(level = relevel(level, "group"))
```

# Plots

```{r, plot-1}
plot <- dat %>%
  ggplot(aes(x    = familiarization_day,
             col   = level,
             shape = level)) +
  geom_point(aes(y = prop),
             size = 3,
             position = position_dodge(width = 0.1)) +
  geom_line(aes(y        = prop,
                group    = level,
                linetype = level),
            position = position_dodge(width = 0.1)) +
  geom_errorbar(aes(ymin = prop - group.sd,
                    ymax = prop + group.sd),
                width = 0.1,
                position = position_dodge(width = 0.1)) +
  scale_color_brewer(palette = "Set1")+
  scale_y_continuous(lim = c(0, 1.05), breaks = seq(0, 1, 0.25)) +
  labs(x = "Familiarization day",
       y = "Proportion of pigs that open a door") +
  my_theme +
  theme(legend.position = "right",
        legend.title = element_blank(),
        legend.text = element_text(size=12))
```

```{r, plot-2, fig.height=10, fig.width=10, warning=FALSE}
plot
```

# How to cite R

"All analyses were performed using R Statistical Software (version 4.2.0; R Core Team 2022)".

Reference: R Core Team (2022). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

```{r, cite-r}
citation()
version$version.string
```

```{r, cite-packages}
citation("tidyverse")
```

# Session Info

```{r, session}
sessionInfo()
```
